---
title: "aquaculture-wc-eez"
format: html
---

# Importing libraries
```{r, message=FALSE}
library(tidyverse)
library(sf)
library(terra)
library(dismo)
library(tmap)
library(here)
```

# Import Data
```{r, message=FALSE, output=FALSE}
# load raster dataset
avg_08 <- rast(here("data","average_annual_sst_2008.tif"))
avg_09 <- rast(here("data","average_annual_sst_2009.tif"))
avg_10 <- rast(here("data","average_annual_sst_2010.tif"))
avg_11 <- rast(here("data","average_annual_sst_2011.tif"))
avg_12 <- rast(here("data","average_annual_sst_2012.tif"))

depth <- rast(here("data","depth.tif"))

# Reading in Shapefile- EEZ Data
regions <- st_read(here("data", "wc_regions_clean.shp"),  quiet = TRUE)


```

### Matching CRS for Rasters
```{r}
# check and transform coordinate reference systems
if(crs(avg_08) == crs(avg_09)) {
  print("Coordinate reference systems match:)")
} else{
  warning("Updating coordinate reference systems to match :(")}

# check and transform coordinate reference systems
if(crs(avg_08) == crs(avg_10)) {
  print("Coordinate reference systems match:)")
} else{
  warning("Updating coordinate reference systems to match :(")}

# check and transform coordinate reference systems
if(crs(avg_08) == crs(avg_11)) {
  print("Coordinate reference systems match:)")
} else{
  warning("Updating coordinate reference systems to match :(")}

# check and transform coordinate reference systems
if(crs(avg_08) == crs(avg_12)) {
  print("Coordinate reference systems match :)")
} else{
  warning("Updating coordinate reference systems to match :(")}

# check and transform coordinate reference systems
if (crs(avg_08) == crs(depth)) {
  print("Coordinate reference systems match :)")
} else {
  warning("Updating coordinate reference systems to match :(")
  # transform raster to match avg_08
  depth <- project(depth, crs(avg_08))
}

# check and transform coordinate reference systems
if (crs(avg_08) == crs(regions)) {
  print("Coordinate reference systems match :)")
} else {
  warning("Updating coordinate reference systems to match :(")
    # transform raster to match avg_08
  regions <- st_transform(regions, crs(avg_08))
}


```
#### Combine SST rasters into a raster stack
```{r}
# Combining rasters
sst_stack <- c(avg_08, avg_09, avg_10, avg_11, avg_12)

# Mean of all Rasters and Converting temp values  K -> °C
sst_mean <- mean(sst_stack) - 273.15
names(sst_mean) <- "sst_mean"

```

#### Checking Crop, Extent, Resolution
```{r}
# Obtain geographic extent/bounding box of species occurrences
st_bbox(sst_mean)
#      xmin       ymin       xmax       ymax 
# -131.98475   29.99305 -114.98672   49.98842 


# Crop depth to the SST extent 
depth_crop <- crop(depth, ext(sst_mean))

# Resample depth to match SST resolution
depth_resampled <- resample(depth_crop, sst_mean, method = "near")

# Check Resolution
if (all(res(sst_mean) == res(depth_resampled))) {
  print("Resolutions matche :)")
} else {
  print("Fix resolution :(")
}

# Check Extent
if (all(ext(sst_mean) == ext(depth_resampled))) {
  print("Extent matches :)")
} else {
  print("Fix Extent :(")
}

# Check CRS
if (crs(sst_mean) == crs(depth_resampled)) {
  print("CRS matches :)")
} else {
  print("Fix CRS :(")
}

```


#Find suitable locations
To find suitable locations for marine aquaculture, we’ll need to find locations that are suitable in terms of both SST and depth.
**Oysters:**
Research has shown that oysters need the following conditions for optimal growth:

- sea surface temperature: 11-30°C
- depth: 0-70 meters below sea level**

- **reclassify SST and depth data into locations that are suitable for oysters*
- hint: set suitable values to 1 and unsuitable values to 0
## find locations that satisfy both SST and depth conditions
```{r}
# Define the reclassification matrix for SST - sea level
rcl_sst <- matrix(c(-Inf, 11, 0, 
                        11, 30, 1,
                        30, Inf, 0), ncol = 3, byrow = TRUE)

# Use classify function
reclassified_sst <- terra::classify(sst_mean, rcl = rcl_sst)


# Define the reclassification matrix for Depth
rcl_depth <- matrix(c(-Inf, -70, 0, 
                        -70, 0, 1,
                        0, Inf, 0), ncol = 3, byrow = TRUE)

# Use classify function
reclassified_depth <- classify(depth_resampled, rcl_depth)

# Combine Oyster and Depth for suitability
suitable <- reclassified_sst * reclassified_depth # Min = 0, Max = 1

# Map it
tm_shape(suitable)+
  tm_raster()
```

## Masking

```{r}
eez_rast <- rasterize(regions, 
                 # template- using extent, resolution, crs
                 suitable, 
                 # use region column
                 field = "rgn")


# Mask
suitable_mask <- mask(suitable,regions)

# Cell size
cell_area <- cellSize(suitable_mask, unit = "km")

#grab area from cell area and mask
suitable_area <- cell_area * suitable_mask

# Sum suitable area by EEZ
area_by_eez_zonal <- zonal(cell_area * suitable_mask, eez_rast, fun = "sum", na.rm = TRUE) %>% 
  rename("suitable_area_km2" = area) %>% 
  mutate(suitable_area_km2 = as.numeric(suitable_area_km2)) %>% 
  as_tibble() %>% 
  arrange(desc(suitable_area_km2))
#print
area_by_eez_zonal
```

```{r, message=FALSE}

tm_shape(regions) +
  tm_polygons("rgn", palette = "Set3", title = "EEZ Regions", alpha = 1) +
tm_shape(suitable_mask) +
  tm_raster(alpha = 0.6, title = "Suitable Area") +
tm_shape(regions) +
  # Adding labels
  tm_text("rgn", size = 0.5) +   
tm_layout(
  legend.outside.position = TRUE)+
tm_scale_bar(position = c(-.01, 0.08)) +
tm_compass(position = c("right", "top"))



```



```{r}
#Set as simple features to conserve geometry
region_sf <- st_as_sf(regions)
# left joing with regions 
#eez_joined <- st_join(region_sf, area_by_eez_zonal, join = 'left')
#convert back to spatvector
#eez_final <- vect(eez_joined, geom= 'geometry' )

# area km2 / suitable area km2 *100 for percentage- table
#eez_final %>% 
  #mutate(perc_area_km2 = area_km2/suitable_area_km2*100)
  


```


### california sea hare
depth range 0 - 22 m
13.4 - 27.1, mean 22 

